Formalizing Stack Safety as a Security Property

### 2023-02-04 submission5/core.tex
CP

% \apt{Too much real estate to sell the generality, flexibility, etc. of our approach.
%   Again, we are not providing evidence that these attributes are true, although we can
%   and should conjecture claim them in passing.}

@@

\begin{itemize}
\item We give a novel characterization of stack safety as a collection
of properties: confidentiality and integrity for callee and caller,
plus well-bracketed control-flow. Different subsets of these
properties characterize the security guarantees offered by mechanisms
that support only some aspects of stack safety\bcp{where do we show
  this?}.  The properties they are parameterized over a notion of
external observation, allowing them to characterize lazy enforcement
mechanisms.
\item We instantiate\bcp{extend??} these definitions in a realistic
setting with features such as\bcp{``such as''? are there others?}
argument passing on the stack, callee-saves registers, and tail-call
elimination.\bcp{How about: We extend these core definitions to
  describe a realistic setting with argument passing on the stack,
  callee-saves registers, and tail-call elimination.}  Our model is
modular enough that handling these features is straightforward.
\bcp{Throughout, the way we talk about the ``core model'' and its
  intantiation / extension / enrichment with realistic features feels
  quite weasily.}
\item We validate a tag-based stack-safety enforcement mechanism,
\emph{Lazy Tagging and Clearing}, via property-based random testing,
find that it falls short of ensuring stack safety, and propose and
validate a fix.
\end{itemize}

@@

When our properties are used to evaluate a system, the threat model will depend on the
details of that system. However, there are some constraints that our design puts on
any system. We must trust that the method of distinguishing security-relevant operations is accurate; if it
involves labels placed on code by a compiler, that means trusting that the compiler placed
those labels correctly. If operations occur that are not recognized, those operations
might not be guaranteed to protect their principals---for instance, an unlabeled call
might not protect the caller's data from the callee. On the flip side, applying an incorrect
label most likely means that the property becomes too strong to be enforced.

Otherwise, we do not assume that the code adheres to any particular
calling convention or implements a source language construct.
In particular, while we are agnostic as to the source
language, we certainly aim to support C, and so any source function might contain undefined
behavior resulting in its compilation to arbitrary machine code. A given enforcement
mechanism or target architecture might place additional constraints, particularly on the behavior of
call and return sequences. For instance, extant implementations tend to assume
implicitly that callee-saved registers have their values maintained by whichever compiler
generated their code. Our properties explicitly state this as a requirement,
which could be enforced by a micro-policy, a well-behaved compiler, or other enforcement technique.
Roessler and DeHon have the compiler in their trusted computing base, so our properties
are stronger than necessary to characterize their micro-policies.

In general, it is impossible to distinguish buggy machine code from an attacker; in
our examples we will identify one function or another as an attacker, but we do not
require any static division between trusted and untrusted code, and we aim to protect
even buggy code.

This is a strong threat model, but it does omit some important aspects of
security enforcement: in particular, hardware and timing attacks are out of scope, and
we do not address concurrency---that is a high-priority future work.
%

\paragraph*{Organization}

In the next section, we will give a brief overview of the structure of properties.
Then in \cref{sec:example}, we will walk through a function
call in a simple example machine and discuss how each of our properties applies to it,
informally. In the process we discuss some motivation for the properties, from a security
perspective.

In \cref{sec:formal} we formalize the machine model,
its {\em security semantics}, and the stack safety properties built on these.
\Cref{sec:enforcement} describes the micro-policies that we test,
\cref{sec:testing} the testing framework itself, and
\cref{sec:relwork,sec:future} discuss related and future work.

>>>>

\apt{Too much real estate to sell the generality, flexibility, etc. of our approach.
  Again, we are not providing evidence that these attributes are true, although we can
  and should conjecture claim them in passing.}

\bcp{Agreed.  I think in particular the first paragraph can just go.
  But this is not enough.  Also, the section is trying to do at least
  two things---talk about earlier work and how we compare, and talk
  about why we believe a flexible approach is good.  The whole
  discussion is a bit wooly and confusing.}

@@

\begin{itemize}
\item We give a novel characterization of stack safety as a collection
of properties: confidentiality and integrity for callee and caller,
plus well-bracketed control-flow. Different subsets of these
properties characterize the security guarantees offered by mechanisms
that support only some aspects of stack safety\bcp{where do we show
  this?}.  The properties they are parameterized over a notion of
external observation, allowing them to characterize lazy enforcement
mechanisms.
\item We instantiate\bcp{extend??} these definitions in a realistic
setting with features such as\bcp{``such as''? are there others?}
argument passing on the stack, callee-saves registers, and tail-call
elimination.\bcp{How about: We extend these core definitions to
  describe a realistic setting with argument passing on the stack,
  callee-saves registers, and tail-call elimination.}  Our model is
modular enough that handling these features is straightforward.
\bcp{Throughout, the way we talk about the ``core model'' and its
  intantiation / extension / enrichment with realistic features feels
  quite weasily.}
\item We validate a tag-based enforcement mechanism, \emph{Lazy
  Tagging and Clearing}, via property-based random testing, find that
it falls short, and propose and validate a fix.
\end{itemize}

In the next section, we give a brief overview of the structure of
properties.  In \cref{sec:example}, we walk through a function call in
a simple example machine and discuss how each of our properties
applies to it, informally. In the process we motivate the properties
from a security perspective.
%
In \cref{sec:formal} we formalize the machine model,
its {\em security semantics}, and the stack safety properties built on these.
In \Cref{sec:enforcement} we describe the micro-policies that we test,
in \cref{sec:testing} the testing framework itself, and in
\cref{sec:relwork,sec:future} related and future work.

@@

When our properties are used to evaluate a system, the threat model
will depend on the details of that system. However, there are some
constraints that our design puts on any system. In particular, we must
trust that the method of distinguishing security-relevant operations
is accurate.  For example, if the enforcement mnechanism involves
labels placed on code by a compiler, we must trust that the compiler
placed those labels correctly. For instance, if a compiled function
call is not marked as such, then the caller's data might not be
protected from the callee.  Conversely, marking too many operations as
calls might mean that the property becomes too strong to be
enforced.\bcp{I don't understand that---wouldn't it mean that
  enforcing the property causes almost all programs to be rejected?
  (I.e., isn't it a case of false negatives rather than false
  positives?)}

\bcp{Still working here:}

We do not assume that low-level code adheres to any fixed calling
convention or implements any particular source-language constructs.
Indeed, if the source
language is C, then high-level programs might contain undefined
behavior, in which case they might be compiled to arbitrary machine
code. A given enforcement 
mechanism or target architecture might place additional constraints, particularly on the behavior of
call and return sequences. For instance, extant implementations tend to assume
implicitly that callee-saved registers have their values maintained by whichever compiler
generated their code. Our properties explicitly state this as a requirement,
which could be enforced by a micro-policy, a well-behaved compiler, or other enforcement technique.
Roessler and DeHon include the compiler in their trusted computing
base, so our properties 
are stronger than necessary to characterize their micro-policies.

In general, it is impossible to distinguish buggy machine code from an
attacker.  In our examples we will identify one function or another as
an attacker, but we do not require any static division between trusted
and untrusted code, and we aim to protect even buggy code.

This is a strong threat model, but it does omit some important aspects
of stack safety in real systems: in particular, it does not address
concurrency.  Hardware and timing attacks are also out of scope.
%
